They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-20 paper/paper.Rmd
better last table

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    affiliations:
      - name: Georgia Institute of Technology
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, I find evidence supprting the hypothesis that round win probability can be modeled as a constant across the series, although not at the naive 50%. Second, I examine post-streak round win probability, in search of evidence positive recency (the "hot hand" fallacy) or negative recency (the "gambler's falacy"). I find that teams perform significantly worse than expected after streaks of 2, 3, and 4 wins when series end up going to 9, 10, or 11 (maximum) rounds, suggesting the presence of negative recency.
bibliography: references.bib  
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(\sigma) = {\binom {r}{\sigma}} \omega^{\sigma}(1-\omega)^{r-\sigma}.
}\label{eq:binom}\end{equation}
```
Plugging in the notional probability $p^{w|kr}_0$ for the null $\omega_0$ and the observed probability $p^{w|kr} = \frac{N^{w|kr}}{P^{w|kr}}$, where $N^{w|kr}$ is the observed number of instances where a teams wins the round immediately following a $k$-round winning streak in a series lasting $r$ rounds, for the observed $\omega$, we can evaluate whether the observed probability is not significantly different from the notional probability. If we reject this hypothesis, then we might consider team momentum, represented by post-streak success, plausible.

@@

```{r}
#| label: cod_round_win_prop_after_k_wins
cod_round_streaks <- cod_rounds |> 
  group_by(series_id, team) |> 
  mutate(
    won_prior_round1 = lag(win_round, n = 1, default = NA),
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA),
    won_prior_round4 = lag(win_round, n = 4, default = NA),
    won_prior_round5 = lag(win_round, n = 5, default = NA)
  ) |> 
  ungroup() |>
  select(
    series_id,
    team,
    is_offense,
    round,
    n_rounds,
    win_series,
    win_round,
    starts_with('won_prior_round')
  )

@@

  cols <- sprintf('won_prior_round%d', 2:k)
  col_syms <- syms(cols)
  cod_round_streaks_after_x <- cod_round_streaks |> 
    drop_na(!!!col_syms) |> 
    count(n_rounds, win_series, win_round, won_prior_round1, !!!col_syms, sort = TRUE)

@@

  suppressMessages(
    cod_round_win_prop_after_z_with_probs <- cod_round_win_prop_after_z |> 
      group_by(win_series, n_rounds) |> 
      mutate(
        total = sum(n),
        prop = n / total
      ) |> 
      ungroup() |> 
      select(n_rounds, win_round, win_series, n, total, prop) |> 
      arrange(n_rounds, win_round) |> 
      inner_join(
        probs_nx_k
      )
  )

  suppressMessages(
    full_join(
      cod_round_win_prop_after_z_with_probs |> 
        nest(data = -c(n_rounds, win_series, win_round)) |> 
        mutate(
          ## TODO: Vectorize this.
          binom_p_value = map_dbl(
            data, 
            ~binom.test(.x$n, .x$total, .x$notional_prop, alternative = 'greater') |> 
              broom::tidy() |>
              pull(p.value)
          ),
          binom_is_signif = binom_p_value <= 0.05
        ) |> 
        unnest(data),
      cod_round_win_prop_after_z_with_probs |> 
        select(n_rounds, win_round, win_series, n, total, prop, notional_prop) |> 
        mutate(
          notional_n = as.integer(n * notional_prop / prop),
          notional_total = as.integer(total * notional_prop / prop), 
          compared_p_value = vectorized_prop_test(n, total, notional_n, notional_total)$p.value,
          compared_is_signif = compared_p_value <= 0.05
        )
    )
  )
}

summarize_cod_win_prop_after_k_wins <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    win_series = c(TRUE, FALSE),
    win_round = c(TRUE, FALSE)
  ) |> 
    mutate(
      is_valid = case_when(
        win_series ~ TRUE,
        (n_rounds - !!k - 6L) >= 0L ~ TRUE,
        TRUE ~ FALSE
      )
    ) |> 
    filter(is_valid) |> 
    select(-is_valid) |> 
    mutate(
      naive_prop = ifelse(win_series, 6L / n_rounds, (n_rounds - 6L) / n_rounds),
      notional_prop = (ifelse(win_series, 6L, n_rounds - 6L) - k) / (n_rounds - k),
      across(c(naive_prop, notional_prop), ~ifelse(win_round, .x, 1 - .x)),
      ms_prop = map2_dbl(
        n_rounds, naive_prop, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      ),
      diff_prop = notional_prop - ms_prop
    ) |> 
    filter(notional_prop > 0, notional_prop < 1)

@@

cod_round_win_prop_after_k_wins <- tibble(win_streak = 3L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
```

```{r}
#| label: cod_round_win_prop_after_k_wins-table
cod_round_win_prop_after_k_wins |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        round(.x, 2) == 1 ~ '1.00',
        TRUE ~ as.character(round(.x, 2))
      )
    ),
    prop_lab = scales::percent(prop, accuracy = 0.1),
    binom_lab = sprintf('%.1f%% (%s)', round(100 * notional_prop, 1), binom_p_value),
    compared_lab = scales::percent(ms_prop, accuracy = 0.1),
    diff_lab = sprintf('%.1f%% (%s)', round(100 * diff_prop, 1), compared_p_value)
  ) |>
  arrange(win_streak, n_rounds, win_series) |> 
  transmute(
    win_series,
    win_round,
    n_rounds,
    total,
    prop_lab,
    binom_lab,
    compared_lab,
    diff_lab
  ) |>
  group_split(win_series, win_round) |> 
  walk(
    function(df) {
      message(sprintf('win_round=%s, win_series=%s', df$win_round[[1]], df$win_series[[1]]))
      df <- df |> 
        select(
          -c(win_series, win_round)
        ) |> 
        gt::gt() |> 
        gt::cols_label(
          # win_streak = 'Streak',
          n_rounds = 'r',
          total = 'N',
          prop_lab = 'O',
          binom_lab = 'a',
          compared_lab = 'b',
          diff_lab = 'c'
        ) |> 
        gt::as_latex() |> 
        cat()
    }
  )

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on streaks of three round wins.[^19] Table \ref{tbl:cod-pw3r-pl3r} shows $P^{w,kr}$ and $P^{l,kr}$ for round win rate and loss rate following $k = 3$ round win streaks respectively.

@@

First,

```{=tex}
\begin{longtable}{rrrrrrr}
\caption{Observed count and proportion of round wins immediately following round win streak of $k=3$, $N^{w|kr}$ and $P^{w|kr}$ respectively, in series lasting $r$ rounds. Additionally, notional expected proportion, $\hat{P}^{w|kr}_0$ and binomial test p-value, assuming notional round probability $p^{w|kr}$ and using observed counts; MS expected proportion $\hat{P}^{w|kr}_{MS}$; and $\hat{D}^{w|kr} = \hat{P}^{w|kr}_{MS} - \hat{P}^{w|kr}_0$ and proportion test p-value. Fields presented in groups, split by whether the team won the round immediately following the 3-round win streak, and whether the team won the series. The post-streak loss analogues are presented as $N^{\ell|kr}$, $P^{\ell|kr}$, $\hat{P}^{\ell|kr}_0$, $\hat{P}^{\ell|kr}_{MS}$, and $\hat{D}^{\ell|kr}$.}
\label{tbl:cod-pw3r-pl3r} \\
\toprule
\multicolumn{6}{c}{\text{win series, win after 3-round win streak}} \\
\midrule
$r$ & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ & $\hat{D}^{w|kr}$ ($p$-value) \\ 
\midrule
7 & 209 & 74.6\% & 75.0\% (0.58) & 75.7\% & -0.7\% (0.94) \\ 
8 & 209 & 62.2\% & 60.0\% (0.28) & 61.9\% & -1.9\% (1.00) \\ 
9 & 193 & 51.8\% & 50.0\% (0.33) & 52.7\% & -2.7\% (0.97) \\ 
10 & 151 & 43.7\% & 42.9\% (0.45) & 44.7\% & -1.8\% (0.94) \\ 
11 & 150 & 40.0\% & 37.5\% (0.29) & 38.8\% & -1.3\% (1.00) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{lose series, win after 3-round win streak}} \\
\midrule
$r$ & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ & $\hat{D}^{w|kr}$ ($p$-value) \\ 
\midrule
10 & 60 & 13.3\% & 14.3\% (0.64) & 26.7\% & -12.4\% (1.00) \\ 
11 & 129 & 24.0\% & 25.0\% (0.63) & 30.8\% & -5.8\% (0.98) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{win series, lose after 3-round win streak}} \\
\midrule
$r$ & $N^{\ell|kr}$ & $P^{\ell|kr}$ & $\hat{P}^{\ell|kr}_0$ ($p$-value) & $\hat{P}^{\ell|kr}_{MS}$ & $\hat{D}^{\ell|kr}$ ($p$-value) \\ 
\midrule
7 & 209 & 25.4\% & 25.0\% (0.48) & 6.4\% & 18.6\% (0.98) \\ 
8 & 209 & 37.8\% & 40.0\% (0.76) & 14.7\% & 25.3\% (0.96) \\ 
9 & 193 & 48.2\% & 50.0\% (0.72) & 22.2\% & 27.8\% (0.97) \\ 
10 & 151 & 56.3\% & 57.1\% (0.62) & 26.7\% & 30.5\% (0.99) \\ 
11 & 150 & 60.0\% & 62.5\% (0.76) & 30.8\% & 31.7\% (0.95) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{lose series, lose after 3-round win streak}} \\
\midrule
$r$ & $N^{\ell|kr}$ & $P^{\ell|kr}$ & $\hat{P}^{\ell|kr}_0$ ($p$-value) & $\hat{P}^{\ell|kr}_{MS}$ & $\hat{D}^{\ell|kr}$ ($p$-value) \\ 
\midrule
10 & 60 & 86.7\% & 85.7\% (0.51) & 44.7\% & 41.1\% (1.00) \\ 
11 & 129 & 76.0\% & 75.0\% (0.45) & 38.8\% & 36.2\% (0.95) \\ 
\bottomrule
\end{longtable}
```
# References {.unnumbered}

>>>>

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    affiliations:
      - name: Georgia Institute of Technology
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, I find evidence supprting the hypothesis that round win probability can be modeled as a constant across the series, although not at the naive 50%. Second, I examine post-streak round win probability, given the series outcome. I find no evidence that post-streak win rate is significantly different than expected for combinations of streak length, series length, and series winner, suggesting that the perception of momentum at the team level is deceiving.
bibliography: references.bib  
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(\sigma) = {\binom {r}{\sigma}} \omega^{\sigma}(1-\omega)^{r-\sigma}.
}\label{eq:binom}\end{equation}
```
Plugging in the notional probability $p^{w|kr}_0$ for the null $\omega_0$ and the observed probability $p^{w|kr} = N^{w|kr} \times P^{w|kr}$, where $N^{w|kr}$ is the observed number of instances where a teams wins the round immediately following a $k$-round winning streak in a series lasting $r$ rounds, for the observed $\omega$, we can evaluate whether the observed probability is not significantly different from the notional probability. If we reject this hypothesis, then we might consider team momentum, represented by post-streak success, plausible.

@@

```{r}
#| label: cod_round_win_prop_after_k_wins
vectorized_binom_test <- function(x, n, p, ...) {
  map_dfr(
    seq_along(x), 
    function(i) {
      res <- binom.test(x[i], n[i], p = p[i], ...)
      tibble(
        estimate = res$estimate,
        conf.low = res$conf.int[[1]],
        conf.high = res$conf.int[[2]],
        p.value = res$p.value
      )
    }
  )
}

cod_round_streaks <- cod_rounds |> 
  group_by(series_id, team) |> 
  mutate(
    won_prior_round1 = lag(win_round, n = 1, default = NA),
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA),
    won_prior_round4 = lag(win_round, n = 4, default = NA),
    won_prior_round5 = lag(win_round, n = 5, default = NA)
  ) |> 
  ungroup() |>
  select(
    series_id,
    team,
    is_offense,
    round,
    n_rounds,
    win_series,
    win_round,
    starts_with('won_prior_round')
  )

@@

  cols <- sprintf('won_prior_round%d', 2:k)
  col_syms <- syms(cols)
  cod_round_streaks_after_x <- cod_round_streaks |> 
    drop_na(!!!col_syms) |> 
    count(n_rounds, win_series, ..., win_round, won_prior_round1, !!!col_syms, sort = TRUE)

@@

  suppressMessages(
    cod_round_win_prop_after_z_with_probs <- cod_round_win_prop_after_z |> 
      group_by(win_series, n_rounds, ...) |> 
      mutate(
        total = sum(n),
        prop = n / total
      ) |> 
      ungroup() |> 
      select(n_rounds, win_round, win_series, ..., n, total, prop) |> 
      # filter(!win_round) |> 
      arrange(n_rounds, win_round, ...) |> 
      inner_join(
        probs_nx_k
      )
  )

  cod_round_win_prop_after_z_with_probs |> 
    mutate(
      notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value,
      ms_p_value = vectorized_binom_test(n, total, p = ms_prop)$p.value
    )
}

summarize_cod_win_prop_after_k_wins <- function(k, ...) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    win_series = c(TRUE, FALSE),
    win_round = c(TRUE, FALSE)
  ) |> 
    mutate(
      is_valid = case_when(
        win_series ~ TRUE,
        (n_rounds - !!k - 6L) >= 0L ~ TRUE,
        TRUE ~ FALSE
      )
    ) |> 
    filter(is_valid) |> 
    select(-is_valid) |> 
    mutate(
      naive_prop = ifelse(win_series, 6L / n_rounds, (n_rounds - 6L) / n_rounds),
      notional_prop = (ifelse(win_series, 6L, n_rounds - 6L) - k) / (n_rounds - k),
      across(c(naive_prop, notional_prop), ~ifelse(win_round, .x, 1 - .x)),
      ms_prop = map2_dbl(
        n_rounds, naive_prop, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      ),
      across(
        ms_prop,
        ~ifelse(win_round, .x, naive_prop + (naive_prop - .x))
      ),
      across(
        ms_prop,
        ~ifelse(is.nan(ms_prop), naive_prop, .x)
      )
    )

@@

summarize_cod_win_prop_after_k_wins_i_rounds <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    round = 3L:11L,
    win_series = c(TRUE, FALSE),
    win_round = c(TRUE, FALSE)
  ) |> 
    filter(round >= (!!k + 1L)) |> 
    mutate(
      is_valid = case_when(
        win_series ~ TRUE,
        (n_rounds - !!k - 6L) >= 0L ~ TRUE,
        TRUE ~ FALSE
      )
    ) |> 
    filter(is_valid) |> 
    select(-is_valid) |> 
    mutate(
      naive_prop = ifelse(win_series, 6L / n_rounds, (n_rounds - 6L) / n_rounds),
      notional_prop = (ifelse(win_series, 6L, n_rounds - 6L) - k) / (n_rounds - k),
      across(c(naive_prop, notional_prop), ~ifelse(win_round, .x, 1 - .x)),
      ms_prop = map2_dbl(
        n_rounds, naive_prop, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      ),
      across(
        ms_prop,
        ~ifelse(win_round, .x, naive_prop + (naive_prop - .x))
      )
    )

  postprocess_round_streaks(
    k = k,
    probs_nx_k = probs_nx_k,
    round
  )
}

cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:5L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value < 0.05)

cod_round_win_prop_after_k_wins_i_rounds <- tibble(win_streak = 2L:3L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins_i_rounds)
  ) |>
  unnest(data)
cod_round_win_prop_after_k_wins_i_rounds |> filter(win_round, notional_p_value < 0.05)
```

```{r}
#| label: cod_round_win_prop_after_k_wins-table
cod_round_win_prop_after_k_wins |> 
  filter(win_streak == 2L) |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    prop_lab = scales::percent(prop, accuracy = 0.1),
    notional_lab = sprintf('%.1f%% (%s)', round(100 * notional_prop, 1), notional_p_value),
    ms_lab = sprintf('%.1f%% (%s)', round(100 * ms_prop, 1), ms_p_value)
  ) |>
  arrange(n_rounds, win_series, win_round) |> 
  transmute(
    n_rounds,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    win_round,
    total,
    prop_lab,
    notional_lab,
    ms_lab
  ) |>
  group_split(win_round) |> 
  walk(
    function(df) {
      message(sprintf('win_round=%s, win_series=%s', df$win_round[[1]], df$win_series[[1]]))
      df <- df |> 
        select(
          -c(win_round)
        ) |> 
        gt::gt() |> 
        gt::cols_label(
          n_rounds = 'r',
          win_series = 'Win series?',
          total = 'N',
          prop_lab = 'O',
          notional_lab = 'a',
          ms_lab = 'b'
        ) |> 
        gt::as_latex() |> 
        cat()
    }
  )

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on streaks of three round wins.[^19] Table \ref{tbl:cod-pw3r-pl3r} compares the observed round win rate $P^{w,kr}$ following streaks of $k=3$ round wins given the series outcome ($r$ rounds and winner) with the notional and MS expected proportions, $\hat{P}^{w|kr}_0$ and $\hat{P}^{w|kr}_{MS}$ respectively.

@@

```{=tex}
\begin{longtable}{rrrrrr}
\caption{Observed count and proportion of round wins immediately following round win streak of $k=3$, $N^{w|kr}$ and $P^{w|kr}$ respectively, given the length of the series ($r$ rounds) and the series winner. Additionally, the notional and MS expected proportions, $\hat{P}^{w|kr}_0$ and $\hat{P}^{w|kr}_{MS}$ respectively, and the $p$-values of binomial tests comparing the underlying notional and MS round win probabilities, $p^{w|kr}_0$ and $p^{w|kr}_{MS}$, to the probability implied by the observed proportion $p^{w|kr} = N^{w|kr} \times P^{w|kr}$.}
\label{tbl:cod-pw3r-pl3r} \\
\toprule
$r$ & \text{Win series?} & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ ($p$-value) \\ 
\midrule
7 & yes & 209 & 74.6\% & 75.0\% (0.94) & 75.7\% (0.69) \\ 
8 & yes & 209 & 62.2\% & 60.0\% (0.53) & 61.9\% (0.94) \\ 
9 & yes & 193 & 51.8\% & 50.0\% (0.67) & 52.7\% (0.83) \\ 
10 & no & 60 & 13.3\% & 14.3\% (1.00) & 26.7\% (0.02) \\ 
10 & yes & 151 & 43.7\% & 42.9\% (0.87) & 44.7\% (0.87) \\ 
11 & no & 129 & 24.0\% & 25.0\% (0.84) & 30.8\% (0.11) \\ 
11 & yes & 150 & 40.0\% & 37.5\% (0.56) & 38.8\% (0.80) \\

\bottomrule
\end{longtable}
```
We observe that, with the exception of $\hat{P}^{w|kr}_{MS}$ when $r = 10$ and the eventual series loser is the team that wins after a streak of three round wins[^20], all binomial test $p$-values are greater than the typical $\alpha = 0.05$ threshold, implying that we cannot reject the null hypothesis that the expected notional and MS post-streak round win rates are different than the observed proportion. As noted before, the MS expectations may be unreliable for CoD SnD, so one is inclined to prefer the results of the notional binomial tests.

[^20]: This implies that the opponent, th eventual series winner, goes on to win the series 6-4 after going down 0-4. This is the only scenario in which one team wins four rounds consecutively but loses a best-of-11 series in 10 rounds.

When performing the same tests for streaks of two, four, and five, there is no case in which I can reject the binomial null hypothesis for the expected notional probability $p^{w|kr}_0$. The null hypothesis can be rejected for expected MS probability $p^{w|kr}_{MS}$ when $k = 4, r = [7, 8, 10]$ and $k = 5, r = [7, 8]$. Nonetheless, the usefulness of $p^{w|kr}_{MS}$ is questionable, as described before, and the sample sizes are smaller for these conditions (all under 20 instances, with the exception of $k = 4, r = [7, 8]$ with 70 and 39 observations respectively).

# References {.unnumbered}

### 2022-08-20 paper/paper.tex
better last table

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, I find evidence
supprting the hypothesis that round win probability can be modeled as a
constant across the series, although not at the naive 50\%. Second, I
examine post-streak round win probability, in search of evidence
positive recency (the ``hot hand'' fallacy) or negative recency (the
``gambler's falacy''). I find that teams perform significantly worse
than expected after streaks of 2, 3, and 4 wins when series end up going
to 9, 10, or 11 (maximum) rounds, suggesting the presence of negative
recency.
\end{abstract}

@@

Plugging in the notional probability \(p^{w|kr}_0\) for the null
\(\omega_0\) and the observed probability
\(p^{w|kr} = \frac{N^{w|kr}}{P^{w|kr}}\), where \(N^{w|kr}\) is the
observed number of instances where a teams wins the round immediately
following a \(k\)-round winning streak in a series lasting \(r\) rounds,
for the observed \(\omega\), we can evaluate whether the observed
probability is not significantly different from the notional
probability. If we reject this hypothesis, then we might consider team
momentum, represented by post-streak success, plausible.

@@

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum (Carlson and Shu 2007), I focus on
streaks of three round wins.\footnote{Three happens to also be a
  reasonable number for series that lasts at maximum 11 rounds.} Table
\ref{tbl:cod-pw3r-pl3r} shows \(P^{w,kr}\) and \(P^{l,kr}\) for round
win rate and loss rate following \(k = 3\) round win streaks
respectively.

First,

\begin{longtable}{rrrrrrr}
\caption{Observed count and proportion of round wins immediately following round win streak of $k=3$, $N^{w|kr}$ and $P^{w|kr}$ respectively, in series lasting $r$ rounds. Additionally, notional expected proportion, $\hat{P}^{w|kr}_0$ and binomial test p-value, assuming notional round probability $p^{w|kr}$ and using observed counts; MS expected proportion $\hat{P}^{w|kr}_{MS}$; and $\hat{D}^{w|kr} = \hat{P}^{w|kr}_{MS} - \hat{P}^{w|kr}_0$ and proportion test p-value. Fields presented in groups, split by whether the team won the round immediately following the 3-round win streak, and whether the team won the series. The post-streak loss analogues are presented as $N^{\ell|kr}$, $P^{\ell|kr}$, $\hat{P}^{\ell|kr}_0$, $\hat{P}^{\ell|kr}_{MS}$, and $\hat{D}^{\ell|kr}$.}
\label{tbl:cod-pw3r-pl3r} \\
\toprule
\multicolumn{6}{c}{\text{win series, win after 3-round win streak}} \\
\midrule
$r$ & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ & $\hat{D}^{w|kr}$ ($p$-value) \\ 
\midrule
7 & 209 & 74.6\% & 75.0\% (0.58) & 75.7\% & -0.7\% (0.94) \\ 
8 & 209 & 62.2\% & 60.0\% (0.28) & 61.9\% & -1.9\% (1.00) \\ 
9 & 193 & 51.8\% & 50.0\% (0.33) & 52.7\% & -2.7\% (0.97) \\ 
10 & 151 & 43.7\% & 42.9\% (0.45) & 44.7\% & -1.8\% (0.94) \\ 
11 & 150 & 40.0\% & 37.5\% (0.29) & 38.8\% & -1.3\% (1.00) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{lose series, win after 3-round win streak}} \\
\midrule
$r$ & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ & $\hat{D}^{w|kr}$ ($p$-value) \\ 
\midrule
10 & 60 & 13.3\% & 14.3\% (0.64) & 26.7\% & -12.4\% (1.00) \\ 
11 & 129 & 24.0\% & 25.0\% (0.63) & 30.8\% & -5.8\% (0.98) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{win series, lose after 3-round win streak}} \\
\midrule
$r$ & $N^{\ell|kr}$ & $P^{\ell|kr}$ & $\hat{P}^{\ell|kr}_0$ ($p$-value) & $\hat{P}^{\ell|kr}_{MS}$ & $\hat{D}^{\ell|kr}$ ($p$-value) \\ 
\midrule
7 & 209 & 25.4\% & 25.0\% (0.48) & 6.4\% & 18.6\% (0.98) \\ 
8 & 209 & 37.8\% & 40.0\% (0.76) & 14.7\% & 25.3\% (0.96) \\ 
9 & 193 & 48.2\% & 50.0\% (0.72) & 22.2\% & 27.8\% (0.97) \\ 
10 & 151 & 56.3\% & 57.1\% (0.62) & 26.7\% & 30.5\% (0.99) \\ 
11 & 150 & 60.0\% & 62.5\% (0.76) & 30.8\% & 31.7\% (0.95) \\ 
\toprule
\toprule
\multicolumn{6}{c}{\text{lose series, lose after 3-round win streak}} \\
\midrule
$r$ & $N^{\ell|kr}$ & $P^{\ell|kr}$ & $\hat{P}^{\ell|kr}_0$ ($p$-value) & $\hat{P}^{\ell|kr}_{MS}$ & $\hat{D}^{\ell|kr}$ ($p$-value) \\ 
\midrule
10 & 60 & 86.7\% & 85.7\% (0.51) & 44.7\% & 41.1\% (1.00) \\ 
11 & 129 & 76.0\% & 75.0\% (0.45) & 38.8\% & 36.2\% (0.95) \\ 
\bottomrule
\end{longtable}

>>>>

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, I find evidence
supprting the hypothesis that round win probability can be modeled as a
constant across the series, although not at the naive 50\%. Second, I
examine post-streak round win probability, given the series outcome. I
find no evidence that post-streak win rate is significantly different
than expected for combinations of streak length, series length, and
series winner, suggesting that the perception of momentum at the team
level is deceiving.
\end{abstract}

@@

Plugging in the notional probability \(p^{w|kr}_0\) for the null
\(\omega_0\) and the observed probability
\(p^{w|kr} = N^{w|kr} \times P^{w|kr}\), where \(N^{w|kr}\) is the
observed number of instances where a teams wins the round immediately
following a \(k\)-round winning streak in a series lasting \(r\) rounds,
for the observed \(\omega\), we can evaluate whether the observed
probability is not significantly different from the notional
probability. If we reject this hypothesis, then we might consider team
momentum, represented by post-streak success, plausible.

@@

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum (Carlson and Shu 2007), I focus on
streaks of three round wins.\footnote{Three happens to also be a
  reasonable number for series that lasts at maximum 11 rounds.} Table
\ref{tbl:cod-pw3r-pl3r} compares the observed round win rate
\(P^{w,kr}\) following streaks of \(k=3\) round wins given the series
outcome (\(r\) rounds and winner) with the notional and MS expected
proportions, \(\hat{P}^{w|kr}_0\) and \(\hat{P}^{w|kr}_{MS}\)
respectively.

\begin{longtable}{rrrrrr}
\caption{Observed count and proportion of round wins immediately following round win streak of $k=3$, $N^{w|kr}$ and $P^{w|kr}$ respectively, given the length of the series ($r$ rounds) and the series winner. Additionally, the notional and MS expected proportions, $\hat{P}^{w|kr}_0$ and $\hat{P}^{w|kr}_{MS}$ respectively, and the $p$-values of binomial tests comparing the underlying notional and MS round win probabilities, $p^{w|kr}_0$ and $p^{w|kr}_{MS}$, to the probability implied by the observed proportion $p^{w|kr} = N^{w|kr} \times P^{w|kr}$.}
\label{tbl:cod-pw3r-pl3r} \\
\toprule
$r$ & \text{Win series?} & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ ($p$-value) \\ 
\midrule
7 & yes & 209 & 74.6\% & 75.0\% (0.94) & 75.7\% (0.69) \\ 
8 & yes & 209 & 62.2\% & 60.0\% (0.53) & 61.9\% (0.94) \\ 
9 & yes & 193 & 51.8\% & 50.0\% (0.67) & 52.7\% (0.83) \\ 
10 & no & 60 & 13.3\% & 14.3\% (1.00) & 26.7\% (0.02) \\ 
10 & yes & 151 & 43.7\% & 42.9\% (0.87) & 44.7\% (0.87) \\ 
11 & no & 129 & 24.0\% & 25.0\% (0.84) & 30.8\% (0.11) \\ 
11 & yes & 150 & 40.0\% & 37.5\% (0.56) & 38.8\% (0.80) \\

\bottomrule
\end{longtable}

We observe that, with the exception of \(\hat{P}^{w|kr}_{MS}\) when
\(r = 10\) and the eventual series loser is the team that wins after a
streak of three round wins\footnote{This implies that the opponent, th
  eventual series winner, goes on to win the series 6-4 after going down
  0-4. This is the only scenario in which one team wins four rounds
  consecutively but loses a best-of-11 series in 10 rounds.}, all
binomial test \(p\)-values are greater than the typical
\(\alpha = 0.05\) threshold, implying that we cannot reject the null
hypothesis that the expected notional and MS post-streak round win rates
are different than the observed proportion. As noted before, the MS
expectations may be unreliable for CoD SnD, so one is inclined to prefer
the results of the notional binomial tests.

When performing the same tests for streaks of two, four, and five, there
is no case in which I can reject the binomial null hypothesis for the
expected notional probability \(p^{w|kr}_0\). The null hypothesis can be
rejected for expected MS probability \(p^{w|kr}_{MS}\) when
\(k = 4, r = [7, 8, 10]\) and \(k = 5, r = [7, 8]\). Nonetheless, the
usefulness of \(p^{w|kr}_{MS}\) is questionable, as described before,
and the sample sizes are smaller for these conditions (all under 20
instances, with the exception of \(k = 4, r = [7, 8]\) with 70 and 39
observations respectively).
